---
title: "PA1_template.Rmd"
output: 
  html_document:
    keep_md: true
---

## LOADING AND PREPROCESSING DATA
Pull out the file activity.csv to read, explore and analyse!!
``` {r read, cache =TRUE, echo = TRUE}
activity <- read.csv("activity.csv")
str(activity)
summary(activity)
```
### Loading the data
``` {r loads, echo = TRUE}
library(plyr)
library(lattice)
library(ggplot2)
```
### Processing and Cleaning the data
``` {r process , echo = TRUE}
activity$day <- weekdays(as.Date(activity$date))
clean <- activity[!is.na(activity$steps), ]
```
## AVERAGE STEPS PER DAY PATTERN
### Creating the aggregate function
``` {r code, echo= TRUE}
sumsteps <- aggregate(activity$steps~activity$date, FUN = sum)
colnames(sumsteps) <- c("dates", "Steps")
head(sumsteps)
```

### Histogram plot
```{r plot1, echo=TRUE}
hist(sumsteps$Steps, ylab = "", xlab = "Total steps", breaks = 5, col = "orange", main = "TOTAL STEPS PER DAY")
abline(v = median(sumsteps$Steps), lty = 2, lwd = 2.5, col = "red")
abline(v = mean(sumsteps$Steps), lty = 3, lwd =2.5, col = "blue")
legend("topright", legend = c("median","mean"), col = c("red","blue"), lty = 2, lwd = 2.5)
```
### Report the mean and median of the dataset
``` {r stats,echo= TRUE}
mean(sumsteps$Steps)
median(sumsteps$Steps)
```
## Time series plot of the average steps taken using ggplot
### Make a ggplot pattern
``` {r timeplot, echo = TRUE}
intervals <- ddply(clean, .(interval), summarize, Avg =  mean(steps))
ggplot(intervals, aes(interval, Avg, col = "red")) + ggtitle("AVERAGE NO.OF STEPS PER DAY") +
  geom_line() + labs(x = "Time Interval", y = "Average steps")
```

### The 5-minute interval that, on average, contains the maximum number of steps???
``` {r max, echo = TRUE}
maximum <- max(intervals$Avg)
intervals[intervals$Avg == maximum, 1]
```
## Impute missing values
### Calculate total number of missing values??
``` {r roe, echo = TRUE}
nrow(activity[is.na(activity$steps), ])
```
My strategy for filling out the NA values is to replace those values with the average steps per day on the basis of interval and day!!
```{r strat,echo= TRUE}
avgsteps <- ddply(clean, .(interval,day), summarize, Avg = mean(steps))
na <- activity[is.na(activity$steps), ]
new <- merge(na, avgsteps, by = c("interval", "day"))
head(new)
new2 <-new[ , c(5,4,1,2)]
colnames(new2) <- c("steps", "date", "interval", "day")
data <- rbind(clean, new2)
head(data)
```
### Histogram plot for imputed and non imputed values
```{r impute, echo=TRUE}
sumsteps2 <- aggregate(data$steps~data$date, FUN = sum)
colnames(sumsteps2) <- c("date","Steps")
mean(sumsteps2$Steps)
median(sumsteps2$Steps)
hist(sumsteps2$Steps, xlab = "steps", ylab = "", main ='IMPUTED AND NON IMPUTED DATA',col = "green")
hist(sumsteps$Steps, xlab = "steps", ylab = "", main ='IMPUTED AND NON IMPUTED DATA',col = "yellow", add = T)
legend("topright", legend = c("imputed data", "non imputed data"), lty = 1, lwd =2.5, col = c("green","yellow"))
```

## Comparing the average number of steps taken per 5-minute interval across weekdays and weekends
``` {r panel, echo= TRUE}
data$daycategory <- ifelse(data$day %in% c("Saturday", "Sunday"), "weekend", "weekday")
intervals <- ddply(data, .(interval,daycategory), summarize, Avg = mean(steps))
head(intervals)
```
### Lattice plot
``` {r lattice, echo= TRUE}
xyplot(Avg~interval|daycategory, data = intervals,type = "l", layout = c(1,2), xlab ="interval", ylab = "steps average", main = "Average steps difference pattern")
```

